---
title: "608 Assignment 1"
author: "James Topor"
date: "January 30, 2017"
output: html_document
---

```{r, echo = FALSE, message=FALSE, warning=FALSE}
# load required packages
library(tidyr, warn.conflicts = FALSE, quietly=TRUE)
library(dplyr, warn.conflicts = FALSE, quietly=TRUE)
library(ggplot2)
```

__1.__ Create a graph that shows the distribution of companies in the dataset by State (ie how many are in each state). There are a lot of States, so consider which axis you should use assuming I am using a portrait oriented screen (ie taller than wide).

```{r, fig.width = 9, fig.height = 7}
inc_df <- read.csv("https://raw.githubusercontent.com/charleyferrari/CUNY_DATA608/master/lecture1/Data/inc5000_data.csv", header = TRUE, stringsAsFactors = FALSE)

# get counts of companies by state
inc_df2 <- data.frame(table(inc_df$State), stringsAsFactors = FALSE)

# reorder by Freq
inc_df2 <- inc_df2[order(inc_df2$Freq, decreasing = TRUE), ]


# display graphic of company counts by state in descending order
ggplot(inc_df2, aes(x= Freq, y=reorder(Var1, Freq))) +
  xlim(0, max(inc_df2$Freq)) +
  geom_segment(aes(yend=Var1), xend=0, colour="grey50") +
  geom_point(size=3, colour = "firebrick") +
  theme_bw() +
  theme(panel.grid.major.y = element_blank()) +
  xlab("Number of Companies per State") +
  ylab("State") +
  ggtitle("Inc. Magazine's 5000 Fastest Growing Companies: Where Are They?")
```


__2.__ Let's dig in on the State with the 3rd most companies in the data set. Imagine you work for the state and are interested in how many people are employed by companies in different industries employ. Create a plot of average employment by industry for companies in this state (only use cases with full data (user R's complete.cases() function). Your graph should show how variable the ranges are, and exclude outliers.

```{r, fig.width = 9, fig.height = 7}
# get all company data for state with 3rd most companies
s3_df <- subset(inc_df, State == inc_df2[3,1])

# get complete cases: doesn't seem to be any incompletes for NY
s3_df <- s3_df[which(complete.cases(s3_df) == TRUE),]


################################################
# Need to exclude outliers and
# generate graphic as horizontal box plot

# get rid of outliers before summarizing
t <- s3_df[!s3_df$Employees %in% boxplot.stats(s3_df$Employees)$out, ]

s4_df <- data.frame(summarise(group_by(t, Industry), 
                    AvgEmps = round(median(as.numeric(Employees)) ) ))

# now order from most to least for graphing
s4_df <- s4_df[order(s4_df$AvgEmps, decreasing = TRUE), ]

# now order from most to least for graphing

t$Industry <- factor(t$Industry, rev(s4_df$Industry))

ggplot(t, aes(x = Industry, y= Employees)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.5, fill = "yellow") +
  theme_bw() + 
  ylab("Median Number of Employees (Outliers Removed)") +
  xlab("Industry") + 
  ylim(0,300) +
  coord_flip() +
  ggtitle("NY: Median Number of Employees per Company per Industry")


```


__3.__ Now imagine you work for an investor and want to see which industries generate the most revenue per employee. Create a chart makes this information clear.

```{r, fig.width = 9, fig.height = 7}

# get complete cases from original data set
rpe_df <- inc_df[which(complete.cases(inc_df) == TRUE),]

# calc rev per employee by industry => group by industry, div rev by num of emps
rpe_df <- data.frame(summarise(group_by(rpe_df, Industry), 
                  RevPerEmp = sum(Revenue) / sum(Employees) / 1000 ))

# now order from most to least for graphing
rpe_df <- rpe_df[order(rpe_df$RevPerEmp, decreasing = TRUE), ]

row.names(rpe_df) <- 1:nrow(rpe_df)
                  
# generate graphic
ggplot(rpe_df, aes(x= RevPerEmp, y=reorder(Industry, RevPerEmp))) +
  xlim(0, max(rpe_df$RevPerEmp)) +
  geom_segment(aes(yend=Industry), xend=0, colour="grey50") +
  geom_point(size=3, colour = "firebrick") +
  theme_bw() +
  theme(panel.grid.major.y = element_blank()) +
  xlab("Revenue per Employee ($Thousands)") +
  ylab("Industry") +
  ggtitle("Which Industries Generate the Most Revenue per Employee?")
```