---
title: "DATA 608 Spring 2017 Final Project: Data Loader"
author: "James Topor"
date: "April 8, 2017"
output: html_document
---

Load libraries
```{r}
options(stringsAsFactors = FALSE)
library(knitr)
library(tidyr, warn.conflicts = FALSE, quietly=TRUE)
library(dplyr, warn.conflicts = FALSE, quietly=TRUE)
library(stringr)
library(WDI)
```

Load WB country name matrix and remove all aggregate regions
```{r}

tmp <- as.data.frame(WDI_data[[2]])

wb_ctry <- as.data.frame(WDI_data[[2]], stringsAsFactors = FALSE)

# remove all 'aggregate' regions since there is no way to
# map them easily
v_countries <- wb_ctry[wb_ctry$region != "Aggregates",]$country

```

Load life expectancy data
```{r}
# load csv file
l_e.raw <- read.csv("https://raw.githubusercontent.com/jtopor/CUNY-MSDA-608/master/FinalProject/DataFiles/life-expectancy.csv", skip=4, header=TRUE, sep = ",", stringsAsFactors = FALSE)

names(l_e.raw)

# create a new data frame using only those columns that match up to TB data
l_exp <- data.frame(l_e.raw$Country.Name,
                    l_e.raw$Country.Code,
                       l_e.raw$X2000,
                       l_e.raw$X2001,
                       l_e.raw$X2002,
                       l_e.raw$X2003,
                       l_e.raw$X2004,
                       l_e.raw$X2005,
                       l_e.raw$X2006,
                       l_e.raw$X2007,
                       l_e.raw$X2008,
                       l_e.raw$X2009,
                       l_e.raw$X2010,
                       l_e.raw$X2011,
                       l_e.raw$X2012,
                       l_e.raw$X2013,
                       l_e.raw$X2014)

# apply meaningful column names that will conform with TB data format
names(l_exp)<-c("country", "code", "x2000",
                           "x2001", "x2002", "x2003", "x2004", "x2005",
                           "x2006", "x2007", "x2008", "x2009", "x2010",
                           "x2011", "x2012", "x2013", "x2014")
# display
kable(head(l_exp), padding=0)

# discard original data frame - no longer needed
rm(l_e.raw)

#################################
# Transform Data to long format:

# gather the 'year' columns
d.df2 <- gather(l_exp, year, life_exp, x2000:x2014, na.rm = FALSE)
kable(head(d.df2))

# clean the "Year" column of the letter 'x'
d.df2$year <- str_extract(d.df2$year, "[[:digit:]]{4}")
kable(head(d.df2))

# sort the data so it is ordered by country then year
d.df2 <- arrange(d.df2, country, year)
kable(head(d.df2))

le.df <- d.df2
# remove l_exp dataframe
rm(l_exp)
```

Load healthcare expenditure data
```{r}
hc.raw <- read.csv("https://raw.githubusercontent.com/jtopor/CUNY-MSDA-608/master/FinalProject/DataFiles/Healthcare-Expenditure.csv", skip=4, header=TRUE, sep = ",", stringsAsFactors = FALSE)

# create a new data frame using only those columns that match up to TB data
hc.df <- data.frame(hc.raw$Country.Name, 
                    hc.raw$Country.Code,
                       hc.raw$X2000,
                       hc.raw$X2001,
                       hc.raw$X2002,
                       hc.raw$X2003,
                       hc.raw$X2004,
                       hc.raw$X2005,
                       hc.raw$X2006,
                       hc.raw$X2007,
                       hc.raw$X2008,
                       hc.raw$X2009,
                       hc.raw$X2010,
                       hc.raw$X2011,
                       hc.raw$X2012,
                       hc.raw$X2013,
                       hc.raw$X2014)

# apply meaningful column names that will conform with TB data format
names(hc.df)<-c("country", "code", "x2000",
                           "x2001", "x2002", "x2003", "x2004", "x2005",
                           "x2006", "x2007", "x2008", "x2009", "x2010",
                           "x2011", "x2012", "x2013", "x2014")
# display
kable(head(hc.df), padding=0)

# discard original data frame - no longer needed
rm(hc.raw)

#################################
# Transform Data to long format:

# gather the 'year' columns
d.df2 <- gather(hc.df, year, hc_exp, x2000:x2014, na.rm = FALSE)
kable(head(d.df2))

# clean the "Year" column of the letter 'x'
d.df2$year <- str_extract(d.df2$year, "[[:digit:]]{4}")
kable(head(d.df2))

# sort the data so it is ordered by country then year
d.df2 <- arrange(d.df2, country, year)
kable(head(d.df2))

hc.df <- d.df2
```

Load GNI data
```{r}
gni.raw <- read.csv("https://raw.githubusercontent.com/jtopor/CUNY-MSDA-608/master/FinalProject/DataFiles/GNI-PerCap.csv", skip=4, header=TRUE, sep = ",", stringsAsFactors = FALSE)

# create a new data frame using only those columns that match up to TB data
gni.df <- data.frame(gni.raw$Country.Name, 
                    gni.raw$Country.Code,
                       gni.raw$X2000,
                       gni.raw$X2001,
                       gni.raw$X2002,
                       gni.raw$X2003,
                       gni.raw$X2004,
                       gni.raw$X2005,
                       gni.raw$X2006,
                       gni.raw$X2007,
                       gni.raw$X2008,
                       gni.raw$X2009,
                       gni.raw$X2010,
                       gni.raw$X2011,
                       gni.raw$X2012,
                       gni.raw$X2013,
                       gni.raw$X2014)

# apply meaningful column names that will conform with TB data format
names(gni.df)<-c("country", "code", "x2000",
                           "x2001", "x2002", "x2003", "x2004", "x2005",
                           "x2006", "x2007", "x2008", "x2009", "x2010",
                           "x2011", "x2012", "x2013", "x2014")
# display
kable(head(gni.df), padding=0)

# discard original data frame - no longer needed
rm(gni.raw)

#################################
# Transform Data to long format:

# gather the 'year' columns
d.df2 <- gather(gni.df, year, percap_gni, x2000:x2014, na.rm = FALSE)
kable(head(d.df2))

# clean the "Year" column of the letter 'x'
d.df2$year <- str_extract(d.df2$year, "[[:digit:]]{4}")
kable(head(d.df2))

# sort the data so it is ordered by country then year
d.df2 <- arrange(d.df2, country, year)
kable(head(d.df2))

gni.df <- d.df2
```

Load TB data
```{r}
tb.raw <- read.csv("https://raw.githubusercontent.com/jtopor/CUNY-MSDA-608/master/FinalProject/DataFiles/TB-per-100K.csv", skip=4, header=TRUE, sep = ",", stringsAsFactors = FALSE)

# create a new data frame using only those columns that match up to TB data
tb.df <- data.frame(tb.raw$Country.Name, 
                    tb.raw$Country.Code,
                       tb.raw$X2000,
                       tb.raw$X2001,
                       tb.raw$X2002,
                       tb.raw$X2003,
                       tb.raw$X2004,
                       tb.raw$X2005,
                       tb.raw$X2006,
                       tb.raw$X2007,
                       tb.raw$X2008,
                       tb.raw$X2009,
                       tb.raw$X2010,
                       tb.raw$X2011,
                       tb.raw$X2012,
                       tb.raw$X2013,
                       tb.raw$X2014)

# apply meaningful column names that will conform with TB data format
names(tb.df)<-c("country", "code", "x2000",
                           "x2001", "x2002", "x2003", "x2004", "x2005",
                           "x2006", "x2007", "x2008", "x2009", "x2010",
                           "x2011", "x2012", "x2013", "x2014")
# display
kable(head(tb.df), padding=0)

# discard original data frame - no longer needed
rm(tb.raw)

#################################
# Transform Data to long format:

# gather the 'year' columns
d.df2 <- gather(tb.df, year, tb_per100K, x2000:x2014, na.rm = FALSE)
kable(head(d.df2))

# clean the "Year" column of the letter 'x'
d.df2$year <- str_extract(d.df2$year, "[[:digit:]]{4}")
kable(head(d.df2))

# sort the data so it is ordered by country then year
d.df2 <- arrange(d.df2, country, year)
kable(head(d.df2))

tb.df <- d.df2

# remove temporary data frame from memory
rm(d.df2)
```


Consolidate all 4 variables into a single data frame and write the file to disk
```{r}
# use one of the 4 variables as the basis for the new consolidated data frame
wb.df <- le.df

wb.df$life_exp <- round(wb.df$life_exp, 2)
# append healthcare exp per capita to new df
wb.df["hc_exp"] <- round(hc.df$hc_exp, 2)
wb.df["gni_percap"] <- gni.df$percap_gni
wb.df["tb_per100K"] <- tb.df$tb_per100K

# check results
kable(head(wb.df))

# now remove all aggregate regions from data frame - should result in 212 unique names
wb.df <- subset(wb.df, country %in% v_countries)

# verify length is 212
length(unique(wb.df$country))

write.csv(wb.df, "c:/SQLData/608/FP/wb2.csv", row.names=FALSE)

# test read of file from Github
# rds_2 <- read.csv("https://raw.githubusercontent.com/jtopor/CUNY-MSDA-608/master/FinalProject/DataFiles/wb2.csv", header=TRUE, sep = ",", stringsAsFactors = FALSE)